---
title: "parseWSI_grades"
author: "Wulcan"
date: "2024-03-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Objective
Parse WSI grades

##Input
-  Human semiquantitative grades (one csv spreadsheet per pathologists) (grades documented in google drive and downloaded as csv) </input/humanSemiQuant_grades/>

- Spreadsheet with slideNames for each image name in the pathologists slides </input/name_image/name.xls>

#### Output
-Parsed semiquantiative grades for intraepithelial lymphocytes <output/wsiAssessment/intraEpithelial_lymphocytes.rds>
-Parsed semiquantiative grades for lamina propria lymphocytes and plasma cells <output/wsiAssessment/laminaPropria_lymphocytesPlasmacells.rds>
- Vector of observers <output/wsiAssessment/observers.rds>

```{r load libraries}
library(tidyverse)
```


### Set batch name and directories
```{r set batch name and directories}
batch <- "test_analysis"

top_dir <- "~/Library/CloudStorage/Box-Box/Projects/AI/paper/AIFeBx_supplemental/" #jmwMacBook
#top_dir <- "C:/Users/jmwulcan/Box/Projects/AI/paper/AIFeBx_supplemental/" #jmwDell
#top_dir <- "C:/Users/15303/Box/Projects/paper/AIFeBx_supplemental/" #JMW lenovo

inputDir_sQuant <- paste0(top_dir,batch, "/input/humanSemiQuant_grades/")

inputDir_AIquant <- paste0(top_dir,batch,"/output/mergeModels/all/slideSummary_2024-03-06.rds")

#Slide names directory
slide_names_dir <- paste0(top_dir, batch, "/input/name_image/names.xls")

#Set output directory 
outputDir_parse <-paste0(top_dir, batch,"/output/wsiGrades_parsed/" )
```

### Parse pathologists semiquantitative grade
Reads in csv files, quality checks, parses and merges
#### Input
-  Human semiquantitative grades (one csv spreadsheet per pathologists) (grades documented in google drive and downloaded as csv) </input/humanSemiQuant_grades/>
- Spreadsheet with slideNames for each image name in the pathologists slides </input/name_image/name.xls>
#### Output
-Parsed semiquantiative grades for intraepithelial lymphocytes <output/wsiAssessment/intraEpithelial_lymphocytes.rds>
-Parsed semiquantiative grades for lamina propria lymphocytes and plasma cells <output/wsiAssessment/laminaPropria_lymphocytesPlasmacells.rds>
- Vector of observers <output/wsiAssessment/observers.rds>

```{r parse semiQuantitative data}

#generate a list of file_paths for csv files in the directory
filePaths_sQuant <- list.files(path = inputDir_sQuant, pattern = "*csv", full.names = TRUE)

#generate a vector of grader initials

graderInitials <- character(0)
fileName <- character(0)
nameParts <- character(0)
graderInitial <- character(0)

for (filePath in filePaths_sQuant) {
  fileName <- basename(filePath)
  nameParts <- strsplit(fileName, "_")[[1]] #separate the elements of the  file path
  graderInitial <- nameParts[3] #the 3rd element is the grader intials
  graderInitial <- gsub(".csv","",graderInitial) #remove .csv from intial
 
  if (!(graderInitial %in% graderInitials)) {
    graderInitials <- c(graderInitials,graderInitial) #generate vector of initials
  }
}

#initate empty list of rawGrades with names from graderInitials
rawGrades <- vector("list", length = length(graderInitials))
names(rawGrades) <- graderInitials


fileName <- character(0)
nameParts <- character(0)
graderInitial <- character(0)
rawGrade<- NULL

#iterate through filePaths and read in rawGrades for each grader
for (initial in graderInitials){
  for (filePath in filePaths_sQuant){
    fileName <- basename(filePath)
    nameParts <- strsplit(fileName, "_")[[1]] #separate the elements of the  file path
    graderInitial <- nameParts[3] #the 3rd element is the grader intials
    graderInitial <- gsub(".csv","",graderInitial)
    
    if(graderInitial == initial) {
      rawGrade <- read.csv(filePath)
    }
    rawGrades[[initial]] <- rawGrade
  }
}

#Quality checks

#Checks if all dataframes adhere to the required structire (105 observations, contains the correct columns and only contains integers between 0 and 3 in numerical columns)

#Check quality criteria for all dataframes in rawGrade
checkedGrades <- vector("list", length = length(graderInitials))
names(checkedGrades) <- graderInitials
checkedGrade <- NULL
checkedGrade_selected <- NULL

for (i in seq_along(rawGrades)){
  checkedGrade <- rawGrades[[i]]
  
  #Check if the dataframe has 105 observations
  if(nrow(checkedGrade) != 105) {
    stop(sprintf("Dataframe %d: Should have 105 observations.", i))
  }
  
  #Check if the dataframe has the correct variables
  required_vars <- c("slide", "lamina_propria_lymphocytes_and_plasmacells", "intraepithelial_lymphocytes")
  
  if (!all(required_vars %in% colnames(checkedGrade))) {
    stop(sprintf("Dataframe %d: should have variables %s.", i, paste(required_vars, collapse = ", ")))
  }
  
  checkedGrade_selected <- checkedGrade %>%
    select(slide, lamina_propria_lymphocytes_and_plasmacells, intraepithelial_lymphocytes)
  
  #Check if the numerical columns are integers between 0 and 3 - if not - round them to the closest (if right in between randomly round up or down)
  
  num_cols <- c("lamina_propria_lymphocytes_and_plasmacells", "intraepithelial_lymphocytes")
  
  if(!all(sapply(checkedGrade_selected[, num_cols], function(x) all(x %in% 0:3) && all(x %% 1 == 0)))) {
    stop(sprintf("Dataframe %d: The numerical columns should contain integers between 0 and 3.", i))
  }
  
  print("All checks passed.")
  
  checkedGrades[[i]] <- checkedGrade_selected
}

#Add prefix and merge

# rename numerical columns p01 to p10 in anticipation of merging

#Function to rename columns and add p01 to p10
rename_and_add_prefix <- function(df, prefix) {
  names(df)[names(df) !=  "slide"] <- paste0(prefix, names(df)[names(df) != "slide"])
  return(df)

}

#Loop through the list of dataframes to rename columns
#observers<- paste0("p",sprintf("%02d",1:length(graderInitials)))

observers <- c("p02", "p05", "p03", "p06", "p11", "p04", "p10", "p09","p08", "p01", "p07") #To match assignments for lc_level validation

rnGrades <- lapply(seq_along(checkedGrades), function(i){
  rename_and_add_prefix(checkedGrades[[i]], paste0(observers[i],"_"))
})

#Sort after pathologist name
roGrades <- NULL
roGrades[["p01"]] <- rnGrades[[10]]
roGrades[["p02"]] <- rnGrades[[1]]
roGrades[["p03"]] <- rnGrades[[3]]
roGrades[["p04"]] <- rnGrades[[6]]
roGrades[["p05"]] <- rnGrades[[2]]
roGrades[["p06"]] <- rnGrades[[4]]
roGrades[["p07"]] <- rnGrades[[11]]
roGrades[["p08"]] <- rnGrades[[9]]
roGrades[["p09"]] <- rnGrades[[8]]
roGrades[["p10"]] <- rnGrades[[7]]
roGrades[["p11"]] <- rnGrades[[5]]


#Merge dataframes from each grader
allGrades <- Reduce(function(x, y) merge(x,y, by="slide", all=TRUE), roGrades)

#Add slide names
slide_names <- read_excel(slide_names_dir)
allGrades <- merge(slide_names, allGrades, by= "slide") 

#Separate IE and LP tables

#Separate up intraepithelial and lamina propria grades in different dataframes

lp_lcPc <- allGrades %>%
  select(slide_name, matches("_lamina_propria_lymphocytes_and_plasmacells")) 

ie_lc <- allGrades %>%
  select(slide_name, matches("_intraepithelial_lymphocytes"))

names(ie_lc) <- gsub("_intraepithelial_lymphocytes","",names(ie_lc))
names(lp_lcPc) <- gsub("_lamina_propria_lymphocytes_and_plasmacells", "", names(lp_lcPc))


  
#Save rds 
saveRDS(ie_lc, paste0(outputDir_sQuant,"WSI_grades_median_intraepithelial.rds"))
saveRDS(lp_lcPc, paste0(outputDir_sQuant,"WSI_grades_median_laminaPropria.rds"))

#Save rds 
saveRDS(ie_lc, paste0(outputDir_parse,"intraEpithelial_lymphocytes.rds"))
saveRDS(lp_lcPc, paste0(outputDir_parse,"laminaPropria_lymphocytesPlasmacells.rds"))
saveRDS(observers, paste0(outputDir_parse,"observers.rds"))
```