---
title: "wsiAI_vs_Path"
author: "Wulcan"
date: "2024-03-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Objective
Compare AI-generated lymphocyte densities to pathologists semiquantitative grades per WSI.

#### Input
-Parsed semiquantiative grades for intraepithelial lymphocytes <output/wsiAssessment/intraEpithelial_lymphocytes.rds>
-Parsed semiquantiative grades for lamina propria lymphocytes and plasma cells <output/wsiAssessment/laminaPropria_lymphocytesPlasmacells.rds>
- AI-derived lymphocyte quantifications </output/mergeModels/all/slideSummary_2024-03-06.rds>

#### Output
**Figure 12** AI-derived lymphocyte counts by semiquantitative grade

```{r load libraries}
library(tidyverse)
options(scipen = 999) #turn of scientific notation
library(ggbeeswarm)
```

### Set batch name and directories
```{r set batch name and directories}

batch <- "test_analysis"

#top_dir <- "~/Library/CloudStorage/Box-Box/Projects/AI/paper/AIFeBx_supplemental/" #jmwMacBook
#top_dir <- "C:/Users/jmwulcan/Box/Projects/AI/paper/AIFeBx_supplemental/" #jmwDell
top_dir <- "C:/Users/15303/Box/Projects/AI/paper/AIFeBx_supplemental/" #JMW lenovo

inputDir_sQuant <- paste0(top_dir, batch,"/output/wsiGrades_parsed/")

inputDir_AIquant <- paste0(top_dir,batch,"/output/mergeModels/all/slideSummary.rds")

#Set output directory 
outputDir_AI_vsPath <-paste0(top_dir, batch,"/output/wsiAI_vs_path/" )

classLabel_palette <- c("Intraepithelial lymphocytes" = "#FA8533","Lamina propria lymphocytes and plasma cells" =  "#3282Bd")

mainFigures_dir <- paste0(top_dir,batch,"/figures/main/")  
```


```{r average}
ie_lc <- readRDS(paste0(inputDir_sQuant,"intraEpithelial_lymphocytes.rds"))
lp_lcPc <- readRDS(paste0(inputDir_sQuant,"laminaPropria_lymphocytesPlasmacells.rds"))


#Add medians
ie_lc <- ie_lc %>%
  rowwise() %>%
  mutate(median_grade = median(c_across(starts_with("p"))),
          class_label = "Intraepithelial lymphocytes")

lp_lcPc <- lp_lcPc %>%
  rowwise() %>%
  mutate(median_grade = median(c_across(starts_with("p"))),
         class_label = "Lamina propria lymphocytes and plasma cells")
      



aiData <- readRDS(inputDir_AIquant)




ie_lc_full <- full_join(ie_lc, aiData, by = "slide_name") %>%
  select(slide_name:class_label, epithelium_lcDensity_1000um2_median, epithelium_lcDensity_1000um2_min, epithelium_lcDensity_1000um2_max, epithelium_lcDensity_1000um2_lower_quantile, epithelium_lcDensity_1000um2_upper_quantile, epithelium_lcDensity_1000um2_range, epithelium_lcDensity_1000um2_iqr) %>%
  rename(density_1000um2_median = epithelium_lcDensity_1000um2_median,
         density_1000um2_min = epithelium_lcDensity_1000um2_min, 
         density_1000um2_max = epithelium_lcDensity_1000um2_max,
         density_1000um2_lower_quantile = epithelium_lcDensity_1000um2_lower_quantile,
         density_1000um2_upper_quantile = epithelium_lcDensity_1000um2_upper_quantile, 
         density_1000um2_range = epithelium_lcDensity_1000um2_range,
         density_1000um2_iqr = epithelium_lcDensity_1000um2_iqr)

lp_lcPc_full <- full_join(lp_lcPc, aiData, by = "slide_name") %>%
  select(slide_name:class_label, lamina_propria_lcDensity_1000um2_median, lamina_propria_lcDensity_1000um2_min, lamina_propria_lcDensity_1000um2_max, lamina_propria_lcDensity_1000um2_lower_quantile, lamina_propria_lcDensity_1000um2_upper_quantile, lamina_propria_lcDensity_1000um2_range, lamina_propria_lcDensity_1000um2_iqr) %>%
  rename(density_1000um2_median = lamina_propria_lcDensity_1000um2_median,
         density_1000um2_min = lamina_propria_lcDensity_1000um2_min, 
         density_1000um2_max = lamina_propria_lcDensity_1000um2_max,
         density_1000um2_lower_quantile = lamina_propria_lcDensity_1000um2_lower_quantile,
         density_1000um2_upper_quantile = lamina_propria_lcDensity_1000um2_upper_quantile, 
         density_1000um2_range = lamina_propria_lcDensity_1000um2_range,
         density_1000um2_iqr = lamina_propria_lcDensity_1000um2_iqr)


wsi_validationData <- rbind(ie_lc_full, lp_lcPc_full) 

#Save rds 
saveRDS(wsi_validationData, paste0(outputDir_AI_vsPath,"WSI_validationData.rds"))


#Plot concordance

plot <- ggplot(wsi_validationData, aes(x = median_grade, y = density_1000um2_median)) +
  geom_beeswarm(alpha = 0.7, aes(color = class_label)) +
  geom_boxplot(aes(group = median_grade), alpha = 0) +
  facet_wrap(~class_label, scales = "free_y") +
  scale_color_manual(values = classLabel_palette, name = "Class" ) +
  scale_x_continuous(breaks = c(0, 1, 2, 3),
                     labels = c("Normal", "Mild", "Moderate", "Marked")) +

 
  labs(
       x = "Semiquantitative grade",
       y = "Lymphocytes/1000um2")+
  theme_bw() +
  theme(legend.position = "none", 
        axis.text = element_text(size = 9, family = "Arial")) 
  
print(plot)

output_file_tif <- paste0(mainFigures_dir,"F12_compare.tif")
tiff(output_file_tif, width = 7, height = 3.5, units = "in", res = 300)
print(plot)
dev.off()


```