---
title: "wsiReproducibility"
author: "Wulcan"
date: "2024-03-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Objective
Calculate interobserver agreement for pathologists semiquantative grades of test set wsis.

##Input
-Parsed semiquantiative grades for intraepithelial lymphocytes <output/wsiAssessment/intraEpithelial_lymphocytes.rds>
-Parsed semiquantiative grades for lamina propria lymphocytes and plasma cells <output/wsiAssessment/laminaPropria_lymphocytesPlasmacells.rds>
- Vector of observers <output/wsiAssessment/observers.rds>

##Output
- **Figure 11 a-b** Upset plots <figures/main/11a_upsetIE.tif><figures/main/11b_upsetLP.tif>
- **Supplemental Figure 9a-b** Heatmap with per slide grades

```{r load libraries}
library(tidyverse)
options(scipen = 999) #turn of scientific notation
library(vcd) # for kappa calculations
library(pheatmap)
library(writexl)
library(UpSetR)
library(RColorBrewer)
```

#Set batch names and directories
```{r set batch names and directories}



batch <- "test_analysis"

top_dir <- "~/Library/CloudStorage/Box-Box/Projects/AI/paper/AIFeBx_supplemental/" #jmwMacBook
#top_dir <- "C:/Users/jmwulcan/Box/Projects/AI/paper/AIFeBx_supplemental/" #jmwDell
#top_dir <- "C:/Users/15303/Box/Projects/paper/AIFeBx_supplemental/" #JMW lenovo

# Set input directory grading files
inputDir <-paste0(top_dir, batch,"/output/wsiGrades_parsed/")

#Set output directory 
outputDir_repro <-paste0(top_dir, batch,"/output/wsiReproducibility/")
supplementaryFigures_dir <- paste0(top_dir,batch,"/figures/supplementary/")
mainFigures_dir <- paste0(top_dir,batch,"/figures/main/")  
tables_dir <- paste0(top_dir,batch,"/tables//")  

classLabel_palette <- c("Intraepithelial lymphocytes" = "#FA8533","Lamina propria lymphocytes and plasma cells" =  "#3282Bd")

grade_labels <- c("0" = "Normal", "1" = "Mild", "2" = "Moderate", "3" = "Marked")

# Define color mapping for each integer value
epithelial_grade_palette <- c("#FDE0C2", "#FDBE85", "#FA8533", "#7F2704")
lamina_propria_grade_palette <- c("#CDE0F1", "#8BBFDC", "#3282Bd", "#08306B")
```


### Upset plots
####Input
-Parsed semiquantiative grades for intraepithelial lymphocytes <output/wsiAssessment/intraEpithelial_lymphocytes.rds>
-Parsed semiquantiative grades for lamina propria lymphocytes and plasma cells <output/wsiAssessment/laminaPropria_lymphocytesPlasmacells.rds>
- Vector of observers <output/wsiAssessment/observers.rds>
####Output
- **Figure 11 a-b** Upset plots
<figures/main/11a_upsetIE.tif><figures/main/11b_upsetLP.tif>
```{r}
#load data 
ie_lc <- readRDS(paste0(inputDir,"intraEpithelial_lymphocytes.rds"))
lp_lcPc <-readRDS(paste0(inputDir,"laminaPropria_lymphocytesPlasmacells.rds"))
observers <- readRDS(paste0(inputDir,"observers.rds"))


ie_upset <- ie_lc %>%
  rowwise() %>%
  mutate(
    Marked = sum(c_across(all_of(observers)) == 3) > 0,
    Moderate = sum(c_across(all_of(observers)) == 2) > 0,
    Mild = sum(c_across(all_of(observers)) == 1) > 0,
    Normal = sum(c_across(all_of(observers)) == 0) > 0
  ) %>%
  select(slide_name, Marked, Moderate, Mild, Normal)




lp_upset <- lp_lcPc %>%
  rowwise() %>%
  mutate(
    Marked = sum(c_across(all_of(observers)) == 3) > 0,
    Moderate = sum(c_across(all_of(observers)) == 2) > 0,
    Mild = sum(c_across(all_of(observers)) == 1) > 0,
    Normal = sum(c_across(all_of(observers)) == 0) > 0
  ) %>%
  select(slide_name, Marked, Moderate, Mild, Normal)



upset_ie <- ie_upset[, c("Marked", "Moderate", "Mild", "Normal")]

upset_ie <- as.data.frame(sapply(upset_ie, as.numeric))

rownames(upset_ie) <- ie_upset$slide_name

ie_upset_plot <- upset(upset_ie, 
                       keep.order = T, 
                       sets = c("Marked", "Moderate", "Mild", "Normal"),
                       main.bar.color = "#FA8533", 
                       sets.bar.color = "#FA8533", c(1.5, 0.9, 0.9, 0.9, 0.9, 0.9))

par(family = "Arial", cex = 0.9)

print(ie_upset_plot)

output_file_tif <- paste0(mainFigures_dir,"F11a_upsetIE.tif")
tiff(output_file_tif, width = 3.5, height = 3.34, units = "in", res = 300)
print(ie_upset_plot)
dev.off()



upset_lp <- lp_upset[, c("Marked", "Moderate", "Mild", "Normal")]

upset_lp <- as.data.frame(sapply(upset_lp, as.numeric))

rownames(upset_lp) <- lp_upset$slide_name

sets_order <- c("Marked", "Moderate", "Mild", "Normal")

lp_upset_plot <- upset(upset_lp, keep.order = T, sets = c("Marked", "Moderate", "Mild", "Normal"), main.bar.color = "#3282Bd", sets.bar.color = "#3282Bd", text.scale = c(0.9, 0.9, 0.9, 0.9, 0.9, 0.9))

par(family = "Arial", cex = 0.9)
 print(lp_upset_plot)

output_file_tif <- paste0(mainFigures_dir,"F11b_upsetLP.tif")
tiff(output_file_tif, width = 3.54, height = 3.34, units = "in", res = 300)
print(lp_upset_plot)
dev.off()

```


### Interobserver agreement
Generate contigency tables, calculate agreement metrics per pathologist pair, plot agreement per pathologist pair, summarize average metrics across pairs. 
##Input
-Parsed semiquantiative grades for intraepithelial lymphocytes <output/wsiAssessment/intraEpithelial_lymphocytes.rds>
-Parsed semiquantiative grades for lamina propria lymphocytes and plasma cells <output/wsiAssessment/laminaPropria_lymphocytesPlasmacells.rds>
- Vector of observers <output/wsiAssessment/observers.rds>

####Output
-Agreement per pathologist pair for intraepithelial lymphocytes as rds   < output/wsiReproducibility/ie_agreementByPair.rds> and excel <tables/ie_agreementByPair.xlsx> 
-Agreement per pathologist pair for lamina propria lymphocytes as rds and excel <tables/lp_agreementByPair.xlsx> 
< output/wsiReproducibility/lp_agreementByPair.rds> 
-**Supplementary figure 10 a and b ** Weighted agreement and Cohen's kappa per pathologist pair <figures/supplementary/S10a_weightedKappa_perPair.tif/>, <figures/supplementary/S10b_cohensKappa_perPair.tif/>,

```{r generate contingency tables}

#load data 
ie_lc <- readRDS(paste0(inputDir,"intraEpithelial_lymphocytes.rds"))
lp_lcPc <-readRDS(paste0(inputDir,"laminaPropria_lymphocytesPlasmacells.rds"))
observers <- readRDS(paste0(inputDir,"observers.rds"))



#Generate contingency tables for each observer pair

ie_contingencyTables <- list()


observer1 <- character(0)
observer2 <- character(0)

ie_contingency_table <- NULL

for (i in 1:(length(observers) - 1)) {
  for (j in (i +1): length(observers)) {
    observer1 <- observers[i]
    observer2 <- observers[j]
    
    #Create a contingency table for observer 1 vs. observer 2
    ie_contingency_table <- xtabs(~ factor(ie_lc[[observer1]], levels = 0:3) + factor(ie_lc[[observer2]], levels = 0:3))
    
    #Store the contingency table in a list
    ie_contingencyTables[[paste0(observer1," vs ",observer2)]] <- ie_contingency_table
  }
}


#For lamina propria
lp_contingencyTables <- list()


observer1 <- character(0)
observer2 <- character(0)

lp_contingency_table <- NULL

for (i in 1:(length(observers) - 1)) {
  for (j in (i +1): length(observers)) {
    observer1 <- observers[i]
    observer2 <- observers[j]
    
    #Create a contingency table for observer 1 vs. observer 2
    lp_contingency_table <- xtabs(~ factor(lp_lcPc[[observer1]], levels = 0:3) + factor(lp_lcPc[[observer2]], levels = 0:3))
    
    #Store the contingency table in a list
    lp_contingencyTables[[paste0(observer1," vs ",observer2)]] <- lp_contingency_table
  }
}


#Generate vector of observer pairs 
observer_pairs <- names(ie_contingencyTables)



#For each pathologist pair, create a table of summary stats (with kappa and CIs)


#Intraepithelial lymphocytes

ie_summaryTables <- list()
contingency_table <- NULL
total_n = NULL
observed_agreement = NULL
K = NULL
CI_K = NULL
summary_table = NULL

for (observer_pair in observer_pairs){
  contingency_table <- ie_contingencyTables[[observer_pair]]
  
  # Calculate total observations
  total_n <- sum(contingency_table)
  
  # Calculate observed agreement
  observed_agreement <- sum(diag(contingency_table)) / total_n
  
  # Calculate cohens kappa & weighted kappa (linear weights) with SE
  K <- Kappa(contingency_table, weight = c("Equal-Spacing", "Fleiss-Cohen"))
  CI_K <- confint(K)
  
  #Create a summary table
  summary_table <- data.frame(
    total_n = total_n,
    observed_agreement = observed_agreement,
    cohens_kappa = K$Unweighted[1],
    cohens_kappa_SE = K$Unweighted[2],
    cohens_kappa_MoE = (CI_K[3]-CI_K[1])/2,
    cohens_kappa_CI_lwr = CI_K[1],
    cohens_kappa_CI_upr = CI_K[3],
    cohens_kappa_sd = K$Unweighted[2] * sqrt(total_n),
    weighted_kappa = K$Weighted[1],
    weighted_kappa_SE = K$Weighted[2],
    weighted_kappa_MoE = (CI_K[4]-CI_K[2])/2,
    weighted_kappa_CI_lwr = CI_K[2],
    weighted_kappa_CI_upr = CI_K[4],
    weighted_kappa_sd = K$Weighted[2] * sqrt(total_n)
    
   
  
  )
  
  ie_summaryTables[[observer_pair]] <- summary_table
}

# Collect in one dataframe
ie_pathPairs <- lapply(seq_along(ie_summaryTables), function(i) {
  path_pair <- paste0(names(ie_summaryTables)[i])
  ie_PathPair <- as.data.frame(ie_summaryTables[[i]])
  ie_PathPair$path_pair <- path_pair
  return(ie_PathPair)
})

ie_summary_frame <- bind_rows(ie_pathPairs)

ie_summary_frame <- ie_summary_frame %>%
  select(path_pair, everything())


#For lamina propria

lp_summaryTables <- list()
contingency_table <- NULL
total_n = NULL
observed_agreement = NULL
K = NULL
CI_K = NULL
summary_table = NULL

for (observer_pair in observer_pairs){
  contingency_table <- lp_contingencyTables[[observer_pair]]
  
  # Calculate total observations
  total_n <- sum(contingency_table)
  
  # Calculate observed agreement
  observed_agreement <- sum(diag(contingency_table)) / total_n
  
  # Calculate cohens kappa & weighted kappa (linear weights) with SE
  K <- Kappa(contingency_table, weight = c("Equal-Spacing", "Fleiss-Cohen"))
  CI_K <- confint(K)
  
  #Create a summary table
  summary_table <- data.frame(
    total_n = total_n,
    observed_agreement = observed_agreement,
    cohens_kappa = K$Unweighted[1],
    cohens_kappa_SE = K$Unweighted[2],
    cohens_kappa_MoE = (CI_K[3]-CI_K[1])/2,
    cohens_kappa_CI_lwr = CI_K[1],
    cohens_kappa_CI_upr = CI_K[3],
    cohens_kappa_sd = K$Unweighted[2] * sqrt(total_n),
    weighted_kappa = K$Weighted[1],
    weighted_kappa_SE = K$Weighted[2],
    weighted_kappa_MoE = (CI_K[4]-CI_K[2])/2,
    weighted_kappa_CI_lwr = CI_K[2],
    weighted_kappa_CI_upr = CI_K[4],
    weighted_kappa_sd = K$Weighted[2] * sqrt(total_n)
    
   
  
  )
  
  lp_summaryTables[[observer_pair]] <- summary_table
}


lp_pathPairs <- lapply(seq_along(lp_summaryTables), function(i) {
  path_pair <- paste0(names(lp_summaryTables)[i])
  lp_PathPair <- as.data.frame(lp_summaryTables[[i]])
  lp_PathPair$path_pair <- path_pair
  return(lp_PathPair)
})

lp_summary_frame <- bind_rows(lp_pathPairs)

lp_summary_frame <- lp_summary_frame %>%
  select(path_pair, everything())

#Add weighted observed agreement

divide_by_total_n <- function(table) {
  total_n <- sum(table)
  return(table/total_n)
}

proportion_tables = list()

for (name in names(ie_contingencyTables)){
  table <- ie_contingencyTables[[name]]
  proportion_table <- divide_by_total_n(table)
  proportion_tables[[name]] <- proportion_table
}

#define weights matrix
weights_matrix <- matrix(c (1, 2/3, 1/3, 0,
                         2/3, 1, 2/3, 1/3,
                         1/3, 2/3, 1, 2/3,
                         0, 1/3, 2/3, 1), nrow = 4, byrow = FALSE)
weighted_tables <- list()

for (name in names(proportion_tables)) {
  table <- proportion_tables[[name]]
  weighted_table <- table * weights_matrix
  weighted_tables[[name]] <- weighted_table
}


# Save the weigthed observed agreement in a list
w_oa_pairs <- data.frame(path_pair = character(0), weighted_observed_agreement = numeric(0))

#Iterate through weighted tables
for (name in names(weighted_tables)) {
  table <- weighted_tables[[name]]
  sum_of_table <- sum(table)
  
  w_oa_pairs <- rbind(w_oa_pairs, data.frame(path_pair = name, weighted_observed_agreement = sum_of_table))
}

ie_summary_frame <- merge(ie_summary_frame, w_oa_pairs, by = "path_pair") %>%
  select(path_pair, total_n, observed_agreement, weighted_observed_agreement, cohens_kappa :weighted_kappa_sd)


#Lamina propria

#Add weighted observed agreement


proportion_tables = list()

for (name in names(lp_contingencyTables)){
  table <- lp_contingencyTables[[name]]
  proportion_table <- divide_by_total_n(table)
  proportion_tables[[name]] <- proportion_table
}

#define weights matrix
weights_matrix <- matrix(c (1, 2/3, 1/3, 0,
                         2/3, 1, 2/3, 1/3,
                         1/3, 2/3, 1, 2/3,
                         0, 1/3, 2/3, 1), nrow = 4, byrow = FALSE)
weighted_tables <- list()

for (name in names(proportion_tables)) {
  table <- proportion_tables[[name]]
  weighted_table <- table * weights_matrix
  weighted_tables[[name]] <- weighted_table
}


# Save the weigthed observed agreement in a list
w_oa_pairs <- data.frame(path_pair = character(0), weighted_observed_agreement = numeric(0))

#Iterate through weighted tables
for (name in names(weighted_tables)) {
  table <- weighted_tables[[name]]
  sum_of_table <- sum(table)
  
  w_oa_pairs <- rbind(w_oa_pairs, data.frame(path_pair = name, weighted_observed_agreement = sum_of_table))
}

lp_summary_frame <- merge(lp_summary_frame, w_oa_pairs, by = "path_pair") %>%
  select(path_pair, total_n, observed_agreement, weighted_observed_agreement, cohens_kappa :weighted_kappa_sd)


saveRDS(ie_summary_frame, paste0(outputDir_repro,"ie_agreementByPair.rds"))
saveRDS(lp_summary_frame, paste0(outputDir_repro,"lp_agreementByPair.rds"))

write_xlsx(ie_summary_frame, paste0(tables_dir,"ie_agreementByPair.xlsx"))
write_xlsx(lp_summary_frame, paste0(tables_dir,"lp_agreementByPair.xlsx"))

#Explore distributions for kappas per pair (can I average)
ie_short <- ie_summary_frame %>% 
  mutate(classLabel = "Intraepithelial lymphocytes") %>%
  select(path_pair, classLabel, cohens_kappa, cohens_kappa_CI_lwr, cohens_kappa_CI_upr, weighted_kappa, weighted_kappa_CI_lwr, weighted_kappa_CI_upr) 

lp_short <- lp_summary_frame %>% 
   mutate(classLabel = "Lamina propria lymphocytes and plasma cells") %>%
  select(path_pair, classLabel, cohens_kappa, cohens_kappa_CI_lwr, cohens_kappa_CI_upr, weighted_kappa, weighted_kappa_CI_lwr, weighted_kappa_CI_upr) 

kappas <- rbind(ie_short, lp_short)

rounded_kappas <- kappas %>%
   mutate(across(where(is.numeric), ~round(., 2)))

# Plot weighted kappas per pathologist pair
plot <- ggplot(rounded_kappas, aes(x = path_pair, y = weighted_kappa, color = classLabel)) +
  geom_point() +
  geom_errorbar(
    aes(ymin = weighted_kappa_CI_lwr, ymax = weighted_kappa_CI_upr),
    width = 0.2,  # Adjust the width of the error bars
    position = position_dodge(0.2)  # Adjust the position of the error bars
  ) +
  labs(x = "Pathologist pair", y = "Weighted kappa") +
  coord_cartesian(ylim = c(-1, 1)) +
  theme_bw() +
  theme(legend.position = "none", 
        axis.text = element_text(size = 11, family = "Arial"),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  facet_wrap(~ classLabel, scales = "free_y", ncol = 1) +
  scale_color_manual(values = classLabel_palette) 
  
print(plot)

# Save the plot as a tif file
output_file_tif <- paste0(supplementaryFigures_dir,"S10a_weightedKappa_perPair.tif")
tiff(output_file_tif, width = 7, height = 3.5, units = "in", res = 300)
print(plot)
dev.off()


#Plot cohens kappa per pathologist pair
plot <- ggplot(rounded_kappas, aes(x = path_pair, y = cohens_kappa, color = classLabel)) +
  geom_point() +
  geom_errorbar(
    aes(ymin = cohens_kappa_CI_lwr, ymax = cohens_kappa_CI_upr),
    width = 0.2,  # Adjust the width of the error bars
    position = position_dodge(0.2)  # Adjust the position of the error bars
  ) +
  labs(x = "Pathologist pair", y = "Cohens kappa") +
  coord_cartesian(ylim = c(-1, 1)) +
  theme_bw() +
  theme(legend.position = "none", 
        axis.text = element_text(size = 11, family = "Arial"),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  facet_wrap(~ classLabel, scales = "free_y", ncol = 1) +
  scale_color_manual(values = classLabel_palette) 
  
print(plot)

# Save the plot as a tif file
output_file_tif <- paste0(supplementaryFigures_dir,"S10b_cohensKappa_perPair.tif")
tiff(output_file_tif, width = 7, height = 3.5, units = "in", res = 300)
print(plot)
dev.off()



#Explore distributions of kappa values per pair
ggplot(rounded_kappas, aes(x = cohens_kappa)) +
  geom_histogram(binwidth = 0.1, fill = "skyblue", color = "black", alpha = 0.7) +
  labs(title = "Histogram of Cohen's Kappa", x = "Cohen's Kappa", y = "Frequency")

ggplot(rounded_kappas, aes(x = weighted_kappa)) +
  geom_histogram(binwidth = 0.1, fill = "skyblue", color = "black", alpha = 0.7) +
  labs(title = "Histogram of Cohen's Kappa", x = "Cohen's Kappa", y = "Frequency")

#Generate summary statistic frame by averaging pair values

# Generate summary statistics for the pathologist pair data
ie_WSIagree_summaryStats <- ie_summary_frame %>%
  summarise(
    n_pathPairs = n(),
    avg_obs_agree = mean(observed_agreement),
    sd_avg_obs_agree = sd(observed_agreement),
    avg_weighted_agree = mean(weighted_observed_agreement),
    sd_avg_weighted_agree = sd(weighted_observed_agreement),
    avg_cohens_kappa = mean(cohens_kappa),
    sd_avg_cohens_kappa = sd(cohens_kappa),
    avg_cohens_kappa_MoE = mean(cohens_kappa_MoE),
    sd_avg_cohens_kappa_MoE = sd(cohens_kappa_MoE),
    avg_weighted_kappa = mean(weighted_kappa),
    sd_avg_weighted_kappa = sd(weighted_kappa),
    avg_weighted_kappa_SE = mean(weighted_kappa_SE),
    sd_weighted_kappa_SE = sd(weighted_kappa_SE),
    avg_weighted_kappa_MoE = mean(weighted_kappa_MoE),
    sd_avg_weighted_kappa_MoE = sd(weighted_kappa_MoE))


# Generate summary statistics for the pathologist pair data
lp_WSIagree_summaryStats <- lp_summary_frame %>%
  summarise(
    n_pathPairs = n(),
    avg_obs_agree = mean(observed_agreement),
    sd_avg_obs_agree = sd(observed_agreement),
    avg_weighted_agree = mean(weighted_observed_agreement),
    sd_avg_weighted_agree = sd(weighted_observed_agreement),
    avg_cohens_kappa = mean(cohens_kappa),
    sd_avg_cohens_kappa = sd(cohens_kappa),
    avg_cohens_kappa_MoE = mean(cohens_kappa_MoE),
    sd_avg_cohens_kappa_MoE = sd(cohens_kappa_MoE),
    avg_weighted_kappa = mean(weighted_kappa),
    sd_avg_weighted_kappa = sd(weighted_kappa),
    avg_weighted_kappa_SE = mean(weighted_kappa_SE),
    sd_weighted_kappa_SE = sd(weighted_kappa_SE),
    avg_weighted_kappa_MoE = mean(weighted_kappa_MoE),
    sd_avg_weighted_kappa_MoE = sd(weighted_kappa_MoE))


#saveRDS(ie_WSIagree_summaryStats, paste0(outputDir_repro,"ie_agreementAverage.rds"))
#saveRDS(lp_summary_frame, paste0(outputDir_repro,"lp_agreementAverage.rds"))

```


### Visualize raw semiquantiative grades
Generates heatmaps of raw semiquantitativ grades
##Input
-Parsed semiquantiative grades for intraepithelial lymphocytes <output/wsiAssessment/intraEpithelial_lymphocytes.rds>
-Parsed semiquantiative grades for lamina propria lymphocytes and plasma cells <output/wsiAssessment/laminaPropria_lymphocytesPlasmacells.rds>
#### Output
-**Supplemental Figure S9a-b** Heatmap of raw grades intraepithelial <figures/supplementary/S9a_rawIeGrades.tif/>, <figures/supplementary/S9b_rawLpGrades.tif/>

```{r}
#Read in intraepithelial scores
ie_lc <- readRDS(paste0(inputDir,"intraEpithelial_lymphocytes.rds"))

column_sums <- colSums(ie_lc[, -1]) 

# Order the pathologists in order of cumulative score (lowest to highest) (to order x-axis)
pathologist_order <- data.frame(
  variable = names(column_sums),
  column_sum = column_sums
)

pathologist_order <- arrange(pathologist_order, column_sum)

ordered_pathologists <- pathologist_order$variable


# Add variable with cumulative per slide score (to order y axis)
ie_lc$sum <- rowSums(ie_lc[, -1])

ie_lc <- ie_lc %>% arrange(desc(sum))

ie_lc_long <- pivot_longer(ie_lc, cols = starts_with("p"), names_to = "Pathologist", values_to = "Grade")

ie_lc_long$slide_name <- factor(ie_lc_long$slide_name, levels = unique(ie_lc$slide_name))


# Plot the tile map
plot <- ggplot(ie_lc_long, aes(x = factor(Pathologist, levels = ordered_pathologists), y = slide_name, fill = factor(Grade))) +
  geom_tile(color = "black") +
  scale_fill_manual(values = epithelial_grade_palette, name = "Grade", labels = c("0" = "Normal", "1" = "Mild", "2" = "Moderate", "3" = "Marked")) +
  theme_bw() +
  theme(legend.position = "right", 
        axis.text = element_text(size = 5, family = "Arial"),
        axis.title = element_text(size = 9, family = "Arial"),
         axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  labs(x = "Pathologist", y = "Slide")

print(plot)

# Save the plot as a tif file
output_file_tif <- paste0(supplementaryFigures_dir,"S9a_rawIeGrades.tif")
tiff(output_file_tif, width = 3.54, height = 6.78, units = "in", res = 300)
print(plot)
dev.off()

#Lamina propria
#load data 
lp_lcPc <-readRDS(paste0(inputDir,"laminaPropria_lymphocytesPlasmacells.rds"))

# Order the pathologists in order of cumulative score (lowest to highest) (to order x-axis)

column_sums <- colSums(lp_lcPc[, -1]) 

pathologist_order <- data.frame(
  variable = names(column_sums),
  column_sum = column_sums
)

pathologist_order <- arrange(pathologist_order, column_sum)

ordered_pathologists <- pathologist_order$variable


lp_lcPc$sum <- rowSums(lp_lcPc[, -1])

lp_lcPc <- lp_lcPc%>% arrange(desc(sum))

lp_lcPc_long <- pivot_longer(lp_lcPc, cols = starts_with("p"), names_to = "Pathologist", values_to = "Grade")

lp_lcPc_long$slide_name <- factor(lp_lcPc_long$slide_name, levels = unique(lp_lcPc$slide_name))


# Plot the tile map
plot <- ggplot(lp_lcPc_long, aes(x = factor(Pathologist, levels = ordered_pathologists), y = slide_name, fill = factor(Grade))) +
  geom_tile(color = "black") +
  scale_fill_manual(values = lamina_propria_grade_palette, name = "Grade", labels = c("0" = "Normal", "1" = "Mild", "2" = "Moderate", "3" = "Marked")) +
  theme_bw() +
  theme(legend.position = "right", 
        axis.text = element_text(size = 5, family = "Arial"),
        axis.title = element_text(size = 9, family = "Arial"),
         axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  labs(x = "Pathologist", y = "Slide")

print(plot)

# Save the plot as a tif file
output_file_tif <- paste0(supplementaryFigures_dir,"S9b_rawLpGrades.tif")
tiff(output_file_tif, width = 3.54, height = 6.78, units = "in", res = 300)
print(plot)
dev.off()
```


